---
title: "Class 17"
author: "Sean Westwood"
format:
  pdf
---
```{r setup, echo=F}
library(knitr)
library(formatR)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Statistical significance: Confidence intervals

## Calculating Confidence intervals

```{r}
transfer <- read.csv("transfer.csv")
```

Difference in means: t.test()

```{r}
t.test(transfer$literate91[transfer$pop82 <= median(transfer$pop82)],
       transfer$literate91[transfer$pop82 >= median(transfer$pop82)])
```

Regression: confint() and summary()

```{r}
model <- lm(literate91~pop82, data = transfer)
confint(model)
summary(model)
```

# Example

```{r}
STAR <- read.csv("STAR.csv")
```

| Name       | Description                                                                                     |
|------------|-------------------------------------------------------------------------------------------------|
| race       | student's race (White = 1, Black = 2, Asian = 3, Hispanic = 4, Native American = 5, Others = 6) |
| classtype  | type of kindergarten class (small = 1, regular = 2, regular with aid = 3)                       |
| g4math     | total scaled score for math portion of fourth grade standardized test                           |
| g4reading  | total scaled score for reading portion of fourth grade standardized test                        |
| yearssmall | number of years in small classes                                                                |
| hsgrad     | high school graduation (did graduate = 1, did not graduate = 0)                                 |

1. We are interested in the difference between reading scores for the small and regular class sizes (excluding regular classes with aide)?  

1.1. What might a hypothesis be?

1.2. What might a null hypothesis be?

1.3. Is the difference significant? 

```{r}
star_sub <- STAR[STAR$classtype == 1 | STAR$classtype == 2,]

t.test(star_sub$g4reading[star_sub$classtype == 2], star_sub$g4reading[star_sub$classtype == 1])

```


2. Repeat question 1, but use math test scores.  Also combine regular and regular with aide classes. 

2.1. What might a hypothesis be?

2.2. What might a null hypothesis be?

2.3. Is the difference significant? 
```{r}

```

3.1. Draft a null hypothesis and a hypothesis for differences that might exist between class size and achievement among minorities. 

3.2. Subset the data to just minorities (non-whites) and then run t.test comparing math scores for small classes with regular classes.  Can you reject the null?

```{r}

```

3.3. Interpret your results.


## Evaluation

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?
4. How long did you spend on the assignment outside of class?
