---
title: "Quiz 1"
author: ""
format: pdf
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

This quiz is based on: Pierotti, Rachel. (2013). "[Increasing Rejection of Intimate Partner Violence: Evidence of Global Cultural Diffusion](http://dx.doi.org/10.1177/0003122413480363)." *American Sociological Review*, 78: 240-265. 

The Demographic and Health Surveys, initiated in 1992, encompass over 300 surveys across developing nations. These surveys feature standardized questions and interviewer training, allowing for cross-country data comparisons. 

The study illustrates that women with greater exposure to global cultural norms—through urban living, secondary education, or media access—are more inclined to reject intimate partner violence. The dataset the following variables.

| Name            | Description                                                             |
|-----------------|-------------------------------------------------------------------------|
| `beat_goesout`  | Percentage of women in each country that think a husband is justified   |
|                 | to beat his wife if she goes out without telling him.                   |
| `beat_burnfood` | Percentage of women in each country that think a husband is justified   |
|                 | to beat his wife if she burns his food.                                 |
| `no_media`      | Percentage of women in each country that rarely encounter a newspaper,  |
|                 | radio or television                                                   |
| `sec_school`    | Percentage of women in each country with secondary or higher education. |
| `year`          | Year of the survey                                                      |
| `region`        | Region of the world                                                     |
| `country`       | Country                                                                 |

```{r}
ipv <- read.csv("ipv.csv")
```

**Round all answers to two decimals**

# Question 1

What is the mean percentage of women who think a husband is justified to beat his wife if she goes out without telling him (beat_goesout) across all countries and years? 

```{r}

```
Answer: 

# Question 2

How many entries (rows) are there in the dataset from the year 2005?

```{r}

```
Answer: 

# Question 3
How many times does the country with the most entries in the dataset appear?

```{r}

```
Answer: 

# Question 4
What is the percentage difference in sec_school between countries in the ‘Latin America’ region and the ‘Sub-Saharan Africa’ region? (Subtract the mean of ‘Sub-Saharan Africa’ from the mean of ‘Latin America’)?

```{r}

```
Answer: 

# Question 5

What is the difference in mean beat_burnfood between the years 2000 and 2010? (Subtract the mean for 2010 from the mean for 2000)

```{r}

```
Answer:

# Question 6

What is the mean percentage of women who think a husband is justified to beat his wife if she burns his food (beat_burnfood) in the ‘Sub-Saharan Africa’ region?

```{r}

```

Answer: