---
title: "Class 14"
author: "Sean Westwood"
format:
  pdf
---
```{r setup, echo=F}
library(knitr)
library(formatR)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

### Plan:
* Multivariate Regression
* Regression and causation

# Multivariate Regression

# Regression and causation: 

Randomization of Treatments Enables Causal Interpretation

Returning to the GOTV experiment from chapter 2:
```{r}
social <- read.csv("social.csv")
```

| Name        | Description                                                                      |
|-------------|----------------------------------------------------------------------------------|
| hhsize      | household size of voter                                                          |
| messages    | GOTV messages voter received (Civic Duty, Control, Neighbors, Hawthorne)         |
| sex         | sex of voter (female or male)                                                    |
| yearofbirth | year of birth of voter                                                           |
| primary2004 | whether a voter voted in the 2004 Primary election (1=voted, 0=abstained)        |
| primary2006 | whether a voter turned out in the 2006 Primary election (1=voted, 0=abstained)   |

1.1. Let's start by estimating the effect of sex and primary voting in 2004 on turnout in the 2006 primary using a regression model.

```{r}

```

1.2. What proportion of women who voted in the 2004 primary would we predict would in the 2006 primary?
```{r}

```

1.3. What proportion of men who voted in the 2004 primary would we predict would in the 2006 primary?

```{r}

```

1.4. Estimate the effects of the message treatments on voting in the 2006 primary. Also save this model to `model.`

```{r}

```

1.5. What is the reference category? Does this make sense?

To change reference categories we use the relevel() function.  It expects two arguments: the variable we are releveling (converted to a factor) and the value of the category we want to set as the new reference.

For comparison let's run the following model:

```{r}
lm(primary2006 ~ sex, data = social)
```

For example, if we wanted to set the reference level for `sex` to "male" we would use the following code.  To show how this changes the output let's run the model from above again.

```{r}
social$sex <- relevel(as.factor(social$sex), ref = "male")
lm(primary2006 ~ sex+messages, data = social)
```

1.6. Relevel `messages` so that "control" is the reference category. Re-run the model from 1.4.  Compare the results to what we got from 1.4.
```{r}

```

1.7. What is the estiamted turnout in the 2006 primary for those in the Hawthorne treatment?
```{r}

```

1.8. What prediction would we make for someone who is a male in a household of 5 who was in the control condition?

```{r}

```

## Evaluation

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?
4. How long did you spend on the assignment outside of class?
