---
title: "Final"
author: ""
format:
  pdf
---

# American National Election Studies

In this exercise, we explore the data of the American National Election Studies (ANES).  The ANES contained survey data on American elections from 1948-2020.  It is designed to measure political attitudes and behaviors, as well as attitudes towards groups of Americans.

We will use a very small subset of the full dataset for this exam.  We will primarily focus on voter traits, voting behavior and attitudes toward specific groups of Americans.

| Variable       | Value                                                                        |
|----------------|------------------------------------------------------------------------------|
| year           | Year of the election                                                         |
| ownhome        | 1 = Yes, 0 = No                                                              |
| gender         | 1 = Man; 0 = Woman                                                           |
| voted          | 1 = voted; 0 = did not vote                                                  |
| race           | 1 = White; 2= Black; 3 = Asian; 4 = American Indian; 5 = Hispanic; 6+ = Other |
| education      | 1 = Grade School or Less; 2 = High School; 3 = Some college; 4 = College +   |
| feeling_jews   | Feeling toward Jews 0-97 (larger numbers are more positive)                  |
| feeling_blacks | Feeling toward blacks 0-97 (larger numbers are more positive)                |
| feeling_white  | Feeling toward whites 0-97 (larger numbers are more positive)                |
| feeling_police | Feeling toward police 0-97 (larger numbers are more positive)                |
| pid            | 1 = Democrats; 2 = Independents; 3 = Republicans                             |




```{r}
anes <- read.csv("anes.csv")
```

*Round all results to two decimal places*

## Question 1

```{r}

```

## Question 2

```{r}

```

## Question 3


```{r}

```

### Question 4


```{r}

```

## Question 5

```{r}

```
## Question 6


```{r}

```
## Question 7

```{r}

```

## Question 8

```{r}

```

## Question 9

```{r}

```

## Question 10

```{r}

```

