---
title: "Class 9"
author: "Sean Westwood"
format:
  pdf
---
```{r setup, echo=F}
library(knitr)
library(formatR)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

### Plan:
* Exam structure
* Practice questions 

## Midterm

* One hour and 50 minutes 
* Mixture of multiple choice and analysis


# Exam Practice

Today we will practice some analysis.  You will work on the practice by yourself.  I will answer questions as if this were an exam.


**Time management will be important during the midterm.  You should be able to answer the questions below in less than 60 minutes**

We will analyze the relationship between indiscriminate violence and insurgent attacks using data about Russian artillery fire in Chechnya from 2000 to 2005. 

This exercise is based on: Lyall, J. 2009. “[Does Indiscriminate Violence Incite Insurgent Attacks?: Evidence from Chechnya.](http://dx.doi.org/10.1177/0022002708330881)” *Journal of Conflict Resolution* 53(3): 331–62.
    
Some believe that indiscriminate violence increases insurgent attacks by creating more cooperative relationships between citizens and insurgents.  Others believe that indiscriminate violence can be effective in suppressing insurgents' activities.  

```{r}
chechen <- read.csv("chechen.csv")
```

| Name       | Description                                                                            |
|------------|----------------------------------------------------------------------------------------|
| village    | Name of Chechnya village                                                               |
| groznyy    | Variable indicating whether a village is in Groznyy (equal to 1) or not (equal to 0)   |
| fire       | Whether Russians struck a village with artillery fire (equal to 1) or not (equal to 0) |
| deaths     | Estimated number of individuals killed during Russian artillery fire                   |
| preattack  | The number of insurgent attacks before Russian artillery fire                          |
| postattack | The number of insurgent attacks after Russian artillery fire                           |

## Question 1
  
1.1. How many unique villages were hit by Russian fire?  Note that villages will appear multiple times (one record for each time the Russians fired).

```{r}

```

1.2. Which village had the largest number of deaths?

```{r}

```

## Question 2

2.1. Did Russian artillery result in a greater number of deaths in Groznyy compared to the villages outside of Groznyy? Compute a mean difference in deaths for the two groups.
```{r}

```

2.1. Conduct the same comparison but use median differences.
```{r}

```

## Question 3
    
3.1. Compare the average number of insurgent attacks after Russian fire for villages hit by artillery fire and those that were not hit. 
```{r}

```

## Question 4

4.1. Compute the mean difference in the `diffattack` (the difference between preattack and postattack) variable between villages shelled and villages not shelled.  
```{r}

```


